Visualizations for the manuscript

### 2022-03-31 metanalysis_visualization/manuscript_Figures.Rmd
determine end date as may 2021 for plots and add consistent x axis for all the plots

```{r}
# cut-off dates used in the analysis, these should not be changed
time_period <-"months" #possible values: days, weeks, months
history_start_date <- as.Date("2019-01-01")
history_end_date <- as.Date("2021-05-31")
pandemic_start_date <- as.Date("2020-03-15")
start_date_plots <- as.Date("2019-01-01")
end_date_plots <- as.Date("2021-05-31")
```

@@

Put all the aggregated counts from the different sites together
```{r}
for( i in 1:length( outputFiles )){
  print(i)
  load(paste0( outputDir, outputFiles[i]))
  if( i == 1 ){
    count_icd_all <- count_icd
    patients_hospitalized_agg_sex_perc_all <- patients_hospitalized_agg_sex_perc
    patient_count_psy_period_psy_all <- patient_count_psy_period_psy
    perc_disorder_group_all <- perc_disorder_group
    ratios_patients_with_without_psy_all <- ratios_patients_with_without_psy
    bootstrapped_fitted_all <- bootstrapped_fitted_df
    bootstrapped_coefficients_all <- bootstrapped_coefficients_df
    length_hospitalisation_values_all <- length_hospitalisation_values
    mental_codes_qc_output_all <- mental_codes_qc_output %>%
      select( disorder_group, description, concept_code, distinct_patients, distinct_observations, siteid )
    table1_all <- table1
  } else {
    count_icd_all <- rbind(count_icd_all, count_icd)
    perc_disorder_group_all <- rbind(perc_disorder_group_all, perc_disorder_group)
    patients_hospitalized_agg_sex_perc_all <- rbind( patients_hospitalized_agg_sex_perc_all, patients_hospitalized_agg_sex_perc)
    mental_codes_qc_output_all <- rbind( mental_codes_qc_output_all,
                                         mental_codes_qc_output %>%
                                           select( disorder_group, description, concept_code, distinct_patients, distinct_observations, siteid )
    )
    table1_all <- rbind( table1_all, table1)
    patient_count_psy_period_psy_all <- rbind( patient_count_psy_period_psy_all,patient_count_psy_period_psy)
    ratios_patients_with_without_psy_all <- rbind( ratios_patients_with_without_psy_all, ratios_patients_with_without_psy)
    bootstrapped_fitted_all <- rbind(bootstrapped_fitted_all, bootstrapped_fitted_df)
    bootstrapped_coefficients_all <- rbind(bootstrapped_coefficients_all, bootstrapped_coefficients_df)
    length_hospitalisation_values_all <- tryCatch(
      rbind(length_hospitalisation_values_all, length_hospitalisation_values),
      error = function(e) length_hospitalisation_values_all
    )
    rm(count_icd,patients_hospitalized_agg_sex_perc,patient_count_psy_period_psy,
       perc_disorder_group,ratios_patients_with_without_psy,
       patient_count_psy_period_psy_clear, bootstrapped_coefficients_df,
       bootstrapped_fitted_df, mental_codes_qc_output, table1, length_hospitalisation_values)
  }
}

@@

@@

```{r Extracting age information}

age_table1_all <- table1_all %>% 
  filter(name == "Age" | (name == "Total" & Category == "Sex")) %>% 
  select(name, siteid, before_pandemic, during_pandemic) %>% 
  pivot_longer(cols = c(before_pandemic, during_pandemic), 
               names_to = "period",
               values_to = "values") %>% 
  pivot_wider(names_from = name, 
              values_from = values, 
              id_cols = c(siteid, period)) %>% 
  mutate(mean_age = as.numeric(sub("\\s\\(.*", "", x = Age, perl = TRUE)),
         sd_age = as.numeric(str_extract(pattern = "(?<=\\(sd:\\s)[0-9]{1,2}\\.[0-9]{1,2}(?=\\))", 
                                         string = Age)),
         Total = as.numeric(sub("\\s\\(.*", "", x = Total, perl = TRUE))
  ) %>% 
  group_by(period) %>% 
  summarise(mean_age = round(sum(Total * mean_age) / sum(Total), 2), 
            sd_age = round(sum(Total * sd_age) / sum(Total), 2)) %>% 
  mutate(Age = paste0(as.character(mean_age), " (sd: ", as.character(sd_age), ")")) %>% 
  select(period, Age) %>% 
  pivot_wider(names_from = period, values_from = Age) %>% 
  mutate(Category = "Age", 
         name = "mean (sd)") %>% 
  select(Category, name, before_pandemic, during_pandemic) %>% 
  rename(before_pandemic_count = before_pandemic, 
         during_pandemic_count = during_pandemic)
```

@@

ggplot(bootstrapped_fitted_all, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
  facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
  geom_line(aes(color = period),
            alpha = 0.03) +
  geom_line(data = quantile_bootstrapped_fitted_all,
            aes(y = value, group = interaction(quantiles, period)),
            linetype = "dashed") +
  geom_line(data = quantile_bootstrapped_fitted_all[quantile_bootstrapped_fitted_all$quantiles == "50%", ],
            aes(y = value, group = period)) +
  geom_point(data = count_icd_all_boot,
             aes(y = percentage, group=NULL),
             alpha = 0.5) +
  scale_colour_manual(values = cbPalette) +
  guides(color = guide_legend(override.aes = list(alpha = 1))) +
  geom_label(data = plot_annotations_df, 
             aes(label=label), 
            x = -Inf,
            y = +Inf,
            hjust = 0,
            vjust = 1,
            size = 4,
            inherit.aes = FALSE) +
  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
       subtitle = "Parametric bootstrap estimation: (1000 iterations)",
       x = "Calendar dates",
       y = "Percentage",
       color = "Period") +
  theme(strip.text = element_text(size = 10),
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 8), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 14),
        plot.subtitle = element_text(size = 12, face = "italic")) 
ggsave("boostrapped_all_extended.png", height = 20, width = 32, units = "cm")
```

```{r}
### Feb 19-April 21
patient_count_psy_period_psy_all <- patient_count_psy_period_psy_all %>%
  left_join( country_map )

ggplot(bootstrapped_fitted_all, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
  facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
  geom_line(aes(color = period),
            alpha = 0.03) +
  geom_line(data = quantile_bootstrapped_fitted_all,
            aes(y = value, group = interaction(quantiles, period)),
            linetype = "dashed") +
  geom_line(data = quantile_bootstrapped_fitted_all[quantile_bootstrapped_fitted_all$quantiles == "50%", ],
            aes(y = value, group = period)) +
  geom_point(data = patient_count_psy_period_psy_all,
             aes(y = percentage, group=NULL),
             alpha = 0.5) +
  scale_colour_manual(values = cbPalette) +
  guides(color = guide_legend(override.aes = list(alpha = 1))) +
  geom_label(data = plot_annotations_df, 
             aes(label=label), 
            x = -Inf,
            y = +Inf,
            hjust = 0,
            vjust = 1,
            size = 4,
            inherit.aes = FALSE) +
  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
       subtitle = "Parametric bootstrap estimation: (1000 iterations)",
       x = "Calendar dates",
       y = "Percentage",
       color = "Period") +
  theme(strip.text = element_text(size = 10),
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 8), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 14),
        plot.subtitle = element_text(size = 12, face = "italic")) 
ggsave("boostrapped_all.png", height = 20, width = 32, units = "cm")
```

@@

for( i in 1:length(sites)){
  print(i)
  subset_site <- patient_count_psy_period_psy_all %>%
    filter( siteid == sites[i])
  ts_seasonal_percent_per_site <- lm(percentage ~ period*time + harmonic(month_year, 1, 12), subset_site)
  p[[i]] <- time_series_analysis_plotting(ts_seasonal_percent_per_site,
                              subset_site,
                              "percentage",
                              clearance_period = FALSE) +
  labs(title = sites[i],
       x = "Calendar dates",
       y = "Percentage") +
     theme(strip.text = element_text(size = 10),
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 8), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 14),
        plot.subtitle = element_text(size = 12, face = "italic"))

@@

counts2 %>% ggplot( aes( x= as.Date( time_p ), y = value, group= group )) +
  geom_point( aes( shape= group, color = group )) +
  facet_wrap2("plotsLabel", axes = "all", scales ="free", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) + 
  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
  labs(x ="", y = "number of patients")+
  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
  scale_color_manual( values = cbPalette ) + 
  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
        strip.text = element_text(size = 10, face = "bold"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA), 
        legend.position="bottom")
```

>>>>

```{r}
# cut-off dates used in the analysis, these should not be changed
time_period <-"months" #possible values: days, weeks, months
#history_start_date <- as.Date("2019-01-01")
#history_end_date <- as.Date("2021-05-31")
pandemic_start_date <- as.Date("2020-03-15")
start_date_plots <- as.Date("2019-01-01")
end_date_plots <- as.Date("2021-05-31")
```

@@

Put all the aggregated counts from the different sites together
```{r}
for( i in 1:length( outputFiles )){
  print(outputFiles[i])
  load(paste0( outputDir, outputFiles[i]))
  if( i == 1 ){
    count_icd_all <- count_icd
    patients_hospitalized_agg_sex_perc_all <- patients_hospitalized_agg_sex_perc
    patient_count_psy_period_psy_all <- patient_count_psy_period_psy
    perc_disorder_group_all <- perc_disorder_group
    ratios_patients_with_without_psy_all <- ratios_patients_with_without_psy
    bootstrapped_fitted_all <- bootstrapped_fitted_df
    bootstrapped_coefficients_all <- bootstrapped_coefficients_df
    length_hospitalisation_values_all <- length_hospitalisation_values
    mental_codes_qc_output_all <- mental_codes_qc_output %>%
      select( disorder_group, description, concept_code, distinct_patients, distinct_observations, siteid )
    table1_all <- table1
  } else {
    count_icd_all <- rbind(count_icd_all, count_icd)
    perc_disorder_group_all <- rbind(perc_disorder_group_all, perc_disorder_group)
    patients_hospitalized_agg_sex_perc_all <- rbind( patients_hospitalized_agg_sex_perc_all, patients_hospitalized_agg_sex_perc)
    mental_codes_qc_output_all <- rbind( mental_codes_qc_output_all,
                                         mental_codes_qc_output %>%
                                           select( disorder_group, description, concept_code, distinct_patients, distinct_observations, siteid )
    )
    table1_all <- rbind( table1_all, table1)
    patient_count_psy_period_psy_all <- rbind( patient_count_psy_period_psy_all,patient_count_psy_period_psy)
    ratios_patients_with_without_psy_all <- rbind( ratios_patients_with_without_psy_all, ratios_patients_with_without_psy)
    bootstrapped_fitted_all <- rbind(bootstrapped_fitted_all, bootstrapped_fitted_df)
    bootstrapped_coefficients_all <- rbind(bootstrapped_coefficients_all, bootstrapped_coefficients_df)
    length_hospitalisation_values_all <- tryCatch(
      rbind(length_hospitalisation_values_all, length_hospitalisation_values),
      error = function(e) length_hospitalisation_values_all
    )
    rm(count_icd,patients_hospitalized_agg_sex_perc,patient_count_psy_period_psy,
       perc_disorder_group,ratios_patients_with_without_psy,
       patient_count_psy_period_psy_clear, bootstrapped_coefficients_df,
       bootstrapped_fitted_df, mental_codes_qc_output, table1, length_hospitalisation_values)
  }
}

@@

# Filter by dates
```{r}
start_date_plots <- as.Date("2019-01-01")
end_date_plots <- as.Date("2021-05-31")

bootstrapped_fitted_all <-  bootstrapped_fitted_all %>%
  filter( time_p <= end_date_plots &
          time_p >= start_date_plots)

count_icd_all <-  count_icd_all %>%
  filter( time_p <= end_date_plots &
          time_p >= start_date_plots)

patient_count_psy_period_psy_all <- patient_count_psy_period_psy_all %>%
  filter( time_p <= end_date_plots &
          time_p >= start_date_plots)

patients_hospitalized_agg_sex_perc_all <- patients_hospitalized_agg_sex_perc_all %>%
  filter( time_p <= end_date_plots &
          time_p >= start_date_plots)

perc_disorder_group_all <- perc_disorder_group_all %>%
  filter( time_p <= end_date_plots &
          time_p >= start_date_plots)

ratios_patients_with_without_psy_all <- ratios_patients_with_without_psy_all %>%  
  filter( time_p <= end_date_plots &
          time_p >= start_date_plots)
```

@@

```{r Extracting age information}
age_table1_all <- table1_all %>% 
  filter(name == "Age" | (name == "Total" & Category == "Sex")) %>% 
  select(name, siteid, before_pandemic, during_pandemic) %>% 
  pivot_longer(cols = c(before_pandemic, during_pandemic), 
               names_to = "period",
               values_to = "values") %>% 
  pivot_wider(names_from = name, 
              values_from = values, 
              id_cols = c(siteid, period)) %>% 
  mutate(mean_age = as.numeric(sub("\\s\\(.*", "", x = Age, perl = TRUE)),
         sd_age = as.numeric(str_extract(pattern = "(?<=\\(sd:\\s)[0-9]{1,2}\\.[0-9]{1,2}(?=\\))", 
                                         string = Age)),
         Total = as.numeric(sub("\\s\\(.*", "", x = Total, perl = TRUE))
  ) %>% 
  group_by(period) %>% 
  summarise(mean_age = round(sum(Total * mean_age) / sum(Total), 2), 
            sd_age = round(sum(Total * sd_age) / sum(Total), 2)) %>% 
  mutate(Age = paste0(as.character(mean_age), " (sd: ", as.character(sd_age), ")")) %>% 
  select(period, Age) %>% 
  pivot_wider(names_from = period, values_from = Age) %>% 
  mutate(Category = "Age", 
         name = "mean (sd)") %>% 
  select(Category, name, before_pandemic, during_pandemic) %>% 
  rename(before_pandemic_count = before_pandemic, 
         during_pandemic_count = during_pandemic)
```

@@

ggplot(bootstrapped_fitted_all, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
  facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
  geom_line(aes(color = period),
            alpha = 0.03) +
  geom_line(data = quantile_bootstrapped_fitted_all,
            aes(y = value, group = interaction(quantiles, period)),
            linetype = "dashed") +
  geom_line(data = quantile_bootstrapped_fitted_all[quantile_bootstrapped_fitted_all$quantiles == "50%", ],
            aes(y = value, group = period)) +
  geom_point(data = count_icd_all_boot,
             aes(y = percentage, group=NULL),
             alpha = 0.5) +
  scale_colour_manual(values = cbPalette) +
  guides(color = guide_legend(override.aes = list(alpha = 1))) +
  geom_label(data = plot_annotations_df, 
             aes(label=label), 
            x = -Inf,
            y = +Inf,
            hjust = 0,
            vjust = 1,
            size = 4,
            inherit.aes = FALSE) +
  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
       subtitle = "Parametric bootstrap estimation: (1000 iterations)",
       x = "Calendar dates",
       y = "Percentage",
       color = "Period") +
  theme(strip.text = element_text(size = 10),
        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
        axis.text.y = element_text(size = 8), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 14),
        plot.subtitle = element_text(size = 12, face = "italic")) 
ggsave("boostrapped_all_extended.png", height = 20, width = 32, units = "cm")
```

@@

for( i in 1:length(sites)){
  print(i)
  subset_site <- patient_count_psy_period_psy_all %>%
    filter( siteid == sites[i])
  ts_seasonal_percent_per_site <- lm(percentage ~ period*time + harmonic(month_year, 1, 12), subset_site)
  p[[i]] <- time_series_analysis_plotting(ts_seasonal_percent_per_site,
                              subset_site,
                              "percentage",
                              clearance_period = FALSE) +
  labs(title = sites[i],
       x = "Calendar dates",
       y = "Percentage") +
      scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
     theme(strip.text = element_text(size = 10),
        axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
        axis.text.y = element_text(size = 8), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 14),
        plot.subtitle = element_text(size = 12, face = "italic"))

@@

counts_plot <- counts2 %>% ggplot( aes( x= as.Date( time_p ), y = value, group= group )) +
  geom_point( aes( shape= group, color = group )) +
  facet_wrap2("plotsLabel", axes = "all", scales ="free", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) + 
  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
  labs(x ="", y = "number of patients")+
  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
  scale_color_manual( values = cbPalette ) + 
  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
        strip.text = element_text(size = 10, face = "bold"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA), 
        legend.position="bottom")
counts_plot
ggsave("counts_plot.png",counts_plot, height = 20, width = 32, units = "cm")
